---
title: "How Green are U.S. Cities compared to the Rest of the World?"
subtitle: "A Comparative Analysis of Urban Greenspace using Stowell et al.'s (2023) global urban greenness indicator dataset for 1,000+ cities (Global Greenspace Indicator Dataset)"
author: "Christopher B."
institute: "cb9110a@american.edu"
date: "2024-03-27"
output:
  beamer_presentation:
    theme: "Pittsburgh"
    colortheme: "orchid"
fontsize: 10pt
classoption: "aspectratio=169"


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)

## Packages
library(tidyverse)
library(knitr)
library(readxl)
library(lubridate)
library(patchwork)
library(readr)
library(tidytext)

## Load data
df = read_csv("greenspace_data_share.csv")

## Set themes
NDVI_Theme = theme_minimal(base_size = 8) +
    theme(
      text = element_text(family = "Times"),
      plot.title = element_text(family = "Times", hjust = 0.5),
      plot.subtitle = element_text(family = "Times", hjust = 0.5),
      plot.background = element_rect(fill = "white", color = NA),
      legend.position = "none",
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.line.x = element_line(color = "black"),
      axis.text.x = element_text(color = "black"),
      axis.text.y = element_text(color = "black")
    )

p2_labs = tibble(
    labels = c('Indonesia','India','USA','Brazil','Nigeria','China')
  )

```

## Background
**Normalized Difference Vegetation Index (NDVI)**
- A widely-used, accurate metric for quantifying the health and density of vegetation using sensor data (e.g., satellites)

**Dataset Contents:**

- Peak and Annual 1x1 km NDVI for over 1,000 urban centers
- Additional information regarding the climate zone
- Level of Development (Human Development Index)


**Uses:**

- Identifying urban centers that lack greenspace
- Developing climate-related plans to counter climate effects in urban centers
- May be adapted to fit other health-related studies



## Global Average NDVI in 2021

```{r Data_Wrangling}
df1 = df |>
  na.omit() |>
  group_by(City) |>
  mutate(
    mean_ndvi_city = mean(annual_avg_2021),
    mean_peak_ndvi_city = mean(peak_NDVI_2021)
    )
df1 = df1 |>
  group_by(Country) |>
  mutate(
    mean_ndvi_country = mean(annual_avg_2021),
    mean_peak_ndvi_country = mean(peak_NDVI_2021)    
  )

df2 = df1 |>
mutate(ann_avg_gradient = case_when(
        mean_ndvi_city < 0.2 ~ "#Ceb010", 
        mean_ndvi_city >= 0.2 & mean_ndvi_city < 0.3 ~ "#C4ce10", 
        mean_ndvi_city >= 0.3 & mean_ndvi_city < 0.4 ~ "#9cce10", 
        mean_ndvi_city >= 0.4 & mean_ndvi_city < 0.5 ~ "#74ce10", 
        mean_ndvi_city >= 0.5 & mean_ndvi_city < 0.6 ~ "#10CE38", 
        mean_ndvi_city >= 0.6 & mean_ndvi_city < 0.7 ~ "#10ce88", 
        mean_ndvi_city >= 0.7 ~ "#10cea3", 
      ),
      peak_gradient = case_when(
        mean_peak_ndvi_city < 0.2 ~ "#Ceb010", 
        mean_peak_ndvi_city >= 0.2 & mean_peak_ndvi_city < 0.3 ~ "#C4ce10", 
        mean_peak_ndvi_city >= 0.3 & mean_peak_ndvi_city < 0.4 ~ "#9cce10", 
        mean_peak_ndvi_city >= 0.4 & mean_peak_ndvi_city < 0.5 ~ "#74ce10", 
        mean_peak_ndvi_city >= 0.5 & mean_peak_ndvi_city < 0.6 ~ "#10CE38", 
        mean_peak_ndvi_city >= 0.6 & mean_peak_ndvi_city < 0.7 ~ "#10ce88", 
        mean_peak_ndvi_city >= 0.7 ~ "#10cea3"
      ),
      grad_peak_NDVI_2021 = case_when(
        peak_NDVI_2021 < 0.2 ~ "#Ceb010", 
        peak_NDVI_2021 >= 0.2 & peak_NDVI_2021 < 0.3 ~ "#C4ce10", 
        peak_NDVI_2021 >= 0.3 & peak_NDVI_2021 < 0.4 ~ "#9cce10", 
        peak_NDVI_2021 >= 0.4 & peak_NDVI_2021 < 0.5 ~ "#74ce10", 
        peak_NDVI_2021 >= 0.5 & peak_NDVI_2021 < 0.6 ~ "#10CE38", 
        peak_NDVI_2021 >= 0.6 & peak_NDVI_2021 < 0.7 ~ "#10ce88", 
        peak_NDVI_2021 >= 0.7 ~ "#10cea3"
      )
  ) #Also tried gradient2 = colorRamp(c("yellow", "green4"))


vis1_text_color = c("red", "#C4ce10", "#9cce10", "#10ce88")

```

```{r summary_stats, include = FALSE}
stats = df2 |>
  select(City, Country, mean_ndvi_city) |>
  group_by(City) |>
  summarize(Country = Country,
            mean_ndvi_city = mean_ndvi_city) |>
  arrange(desc(mean_ndvi_city))
head(stats)

vis1_quant_25 <- quantile(df2$mean_ndvi_city, 0.25)
vis1_quant_75 <- quantile(df2$mean_ndvi_city, 0.75)

city_quantile_25 <- df2$City[which.min(abs(df2$mean_ndvi_city - vis1_quant_25))]
city_quantile_75 <- df2$City[which.min(abs(df2$mean_ndvi_city - vis1_quant_75))]
```

```{r Plot 1, fig.dim=c(5,2.5)}
p1 = df2 |>
    ggplot(aes(
      x = mean_ndvi_city
    )) +
    geom_boxplot(
        fill = "grey90",
        outliers = TRUE,
        outlier.color = "green4",
        outlier.fill = "green4",
        outlier.shape = 19,
        outlier.size = 1.5,
        outlier.stroke = 0.5,
        outlier.alpha = 0.3,
        notch = FALSE,
        notchwidth = 0.75,
        varwidth = TRUE,
        show.legend = NA
    ) + 
    labs(
      x = "Annual NDVI as mean of seasonal NDVI values",
      title = "Global Annual NDVI (City-Level)",
      subtitle = "How the Average U.S. City Compares with Global Mean Annual NDVI"
    ) +
    scale_x_continuous(
      name = "Annual NDVI as mean of seasonal NDVI values", 
      breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6),
      labels = c('0', 'Exceptionally Low', 'Very Low', 'Low', 'Moderate', 'High', 'Very High')
    ) +
    scale_y_continuous(
      name = NULL,
      breaks = NULL,
      labels = NULL
    ) +
  geom_segment(
    x = 0.3185625, #this red line shows the average U.S. city's mean NDVI
    y = -5,
    xend = 0.3185625,
    yend = 5,
    color = "red",
    linewidth = 1
  ) +
  geom_text(aes(x = 0, y = 0.3185625, label = "Average U.S. City (0.319)"), color = "red", size = 2.5, hjust = -2.375, vjust = 2, family = "Times") + #annotation for red line
  
  geom_text(aes(x = 0, y = 0.4, label = "Kandy, Sri Lanka (0.632)"), color = "#10ce88", size = 2.5, hjust = -3.5, vjust = 24, family = "Times") +
  geom_segment(aes(x = 0.65, xend = 0.633, y = -0.28, yend = -0.02), arrow = arrow(length = unit(0.1, "inches")), color = "#10ce88") + #arrow for Kandy, Sri Lanka
  
  geom_text(aes(x = 0, y = 0.4, label = "Curitiba, Brazil (0.338)"), color = "#9cce10", size = 2.5, hjust = -2.65, vjust = 24, family = "Times") +
  geom_segment(aes(x = 0.38, xend = 0.34, y = -0.28, yend = -0.02), arrow = arrow(length = unit(0.1, "inches")), color = "#9cce10") +
  
  geom_text(aes(x = 0, y = 0.4, label = "Kaifeng, China (0.227)"), color = "#C4ce10", size = 2.5, hjust = -0.7, vjust = 24, family = "Times") +
  geom_segment(aes(x = 0.2, xend = 0.225, y = -0.28, yend = -0.02), arrow = arrow(length = unit(0.1, "inches")), color = "#C4ce10") +
  
  NDVI_Theme
p1
```





## Peak NDVI


``` {r Plot 2, fig.dim=c(4.5,2.5)}

p2 = df2 |>
    group_by(Country, mean_peak_ndvi_country) |> 
    filter(
      Country == "United States of America" |
      Country == "China" |
      Country == "India" |
      Country == "Indonesia" |
      Country == "Nigeria" |
      Country == "Russia" |
      Country == "Brazil"
    ) |> #tried to get a wide span of countries

    ggplot(aes(
      x = Country,
      y = mean_peak_ndvi_country,
      fill = Country
    )) +
    labs(
      title = "Peak NDVI per Country",
      x = NULL,
      y = "Mean Peak NDVI"
    ) +
    geom_col(
        aes(x = reorder(Country, mean_peak_ndvi_country)),
        position = position_dodge(),
        stat = "identity",
        show.legend = FALSE
    ) +
    scale_fill_manual(
      values = c("#9cce10", "#9cce10", "#9cce10", "#74ce10", "#9cce10", "#9cce10")
      #Original: values = c("green", "red3", "orange", "red", "#008000", "blue")
    ) +
    coord_flip() +
    NDVI_Theme +
    theme(panel.grid = element_blank(),
         panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank())


df3 = df2 |>
    group_by(Country, peak_NDVI_2021) |> 
    filter(
      Country == "United States of America"
    )

p3 = df2 |>
    group_by(Country, peak_NDVI_2021) |> 
    filter(
      Country == "United States of America"
    ) |> 
    ggplot(aes(
      x = City,
      y = peak_NDVI_2021,
      fill = grad_peak_NDVI_2021
    )) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + #fixes the 'ghostly hover' issue where bar chart hovers above the x-axis
  scale_fill_manual(values = c(
      "#Ceb010" = "#Ceb010",
      "#C4ce10" = "#C4ce10",
      "#9cce10" = "#9cce10",
      "#74ce10" = "#74ce10",
      "#10CE38" = "#10CE38",
      "#10ce88" = "#10ce88",
      "#10cea3" = "#10cea3"
    )) +
    labs(
      title = "Peak NDVI across U.S. Cities",
      x = "U.S. Cities",
      y = "Mean Peak NDVI"
    ) +
    geom_col(
        aes(x = reorder(City, -peak_NDVI_2021)),
        position = position_dodge(),
        stat = "identity",
        show.legend = TRUE
    ) +
    scale_x_discrete(
      labels = NULL
    ) +
    coord_flip() +
    NDVI_Theme +
    theme(panel.grid = element_blank(),
         panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank())


p2 + p3 +
     plot_annotation(
       title = 'Peak NDVI in Select Countries and U.S. Cities',
       theme = theme(plot.title = element_text(size = 10, family = "Times")))
```
